The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-07 manuscript/manuscript.Rnw
discuss pig and laurasiatheria specific results

@@

\begin{table}[h]
  \tiny
<<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
top.pig.genes <- top.genes.table(species.name="sus scrofa")
print(xtable(top.pig.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.pig.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \pig}
\label{tab:top50_pig_placenta}
\end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
\cref{tab:top50_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}.

>>>>

\begin{table}
  \tiny
<<lau.lineage.specific,echo=FALSE,results="asis">>=
lau.all.lin.spec <- placenta.lineage.specific.table("lau.all")
setkey(lau.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed in the Laurasiatheria
  superorder as compared to all other analyzed species.}
\label{tab:laurasiatheria_lineage_specific}
\end{table}

\begin{table}
  \tiny
<<eua.lau.lineage.specific,echo=FALSE,results="asis">>=
eua.lau.lin.spec <- placenta.lineage.specific.table("eua.lau")
setkey(eua.lau.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed between the
  Euasiatherian and Laurasiatherian superorder.}
\label{tab:laurasiatheria_eurasiatherian_lineage_specific}
\end{table}

We examined the expression of five Laurasiatherians: \cow, \sheep,
\pig, \horse, and \dog. We were unable to detect any significant
differences after correcting for multiple testing between
Laurasiatherians and all other analyzed species
(\cref{tab:laurasiatheria_lineage_specific}) or between
Laurasiatherians and Euasiatherians
(\cref{tab:laurasiatheria_eurasiatherian_lineage_specific}).

@@

\begin{table}[h]
  \tiny
<<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
top.pig.genes <- top.genes.table(species.name="sus scrofa")
print(xtable(top.pig.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.pig.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\pig}{tab:all_sus_scrofa_placenta}}
\label{tab:top50_pig_placenta}
\end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
\cref{tab:top50_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}. The top two
expressed transcripts,
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000025162",complete=TRUE)}
and
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000021356",complete=TRUE)},
are miRNA which correspond to the \cow miRNA
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MI0013080}{bta-mir-2904-1}
and
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MIMAT0013845}{bta-mir-2887-1},
respectively, both of which are also expressed in \cow ovary
\cite{Hossain.ea2009:Identificationandcharacterizationof}, although
neither of which have an obvious target gene on the basis of
\textit{in silico} analysis
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}. Other highly
expressed genes include members of the insulin growth factor family
(\Sexpr{gene.expression("sus_scrofa","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("sus_scrofa","IGFBP3",complete=TRUE)}), members
of the annexin complex
(\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","ANXA2",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A12",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A10",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)}), the
galectin \Sexpr{gene.expression("sus_scrofa","LGALS3",complete=TRUE)}
and multiple cytoskeleton and extracellular matrix proteins
(\Sexpr{gene.expression("sus_scrofa","KRT19",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","COL1A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","BETA-ACTIN",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","CLDN4",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)}).
